Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-05-07 manuscript.Rmd
Updates for revision

Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.

@@

Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect]. \Revision{The integral is used as an approximation of the sum for continuous parameters to calculate the probability density.} An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, an estimation of the general and individual effects cannot be provided. However, even proper priors can result in improper posteriors [@hobert1996effect]. Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of this problem, this can result in misleading conclusions. Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors. There is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.

@@

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Student's t}(3, 1, 10), (\#eq:chosenpriors1) \\ 
\mu_{\beta}, \mu_{\delta} &\sim \text{Normal}(0, 1), \\ 
\sigma, \sigma_{\gamma}, \sigma_{\beta}, \sigma_{\delta} &\sim \text{Truncated-Student's t}_{+}(3, 0, 10). 
\end{aligned}
\end{equation}
The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances. Prior distributions on variances have a smaller scale than priors on standard deviation. Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself. In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation. For the `brms` priors, we obtain the following implied prior distributions on variances:

@@

\Revision{## Technical Possibilities are Endless}

>>>>

Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena. We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al. [-@gelman2020] and Schad et al.[-@schad2019toward; -@schad2021workflow], we don't shy away from in-depth explanations for the most curious scientists. To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations. However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}

@@

Another important issue is that wide priors are often improper probability distributions [@hobert1996effect]. \Revision{A prior is proper if two conditions hold: 1. All values of the probability distribution are positive definite, and 2. the probability distribution sums to one for discrete data or integrates to one for continuous data. An example of a proper and improper probability distribution are provided in Figure \@ref(fig:properfigure). The figure shows two hypothetical priors for parameter $\theta$. The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4). To assess the second condition, we need to sum up the function values. Panel A shows a proper distribution where the values sum to 1; panel B shows an improper distribution where the values sum to 1.2.}

```{r properfigure, echo = F, fig.width=7, fig.height=3.5, fig.align='center', fig.cap="Two hypothetical priors for a discrete parameter $\\theta$. A. A proper probability distribution that sums to one. B. An improper probability distribution that sums to 1.2."}
layout(matrix(1:2, nrow = 1))
par(mgp = c(2, .7, 0))
x <- 1:5
y <- c(12, 6, 6, 3, 3)
y <- y / sum(y)

plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), ylab = "Probability", xlab = expression(theta), main = "A. Proper", xlim = c(.5, 5.5))
axis(1, x)
axis(2, c(0, unique(y), 1), las = 2)
for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)

par(mgp = c(2, .7, 0))
y[3] <- .4

plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), xlim = c(.5, 5.5), ylab = "Probability", xlab = expression(theta), main = "B. Improper")
axis(1, x)
axis(2, c(0, unique(y), 1), las = 2)
for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
```

A \Revision{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, \Revision{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors [@hobert1996effect]. Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of the impropriety problem, this can result in misleading conclusions. Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors. \Revision{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below. \Revision{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}

@@

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Student's t}(3, 1, 10), (\#eq:chosenpriors1) \\ 
\mu_{\beta}, \mu_{\delta} &\sim \text{Normal}(0, 1), \\ 
\sigma, \sigma_{\gamma}, \sigma_{\beta}, \sigma_{\delta} &\sim \text{Truncated-Student's t}_{+}(3, 0, 10). 
\end{aligned}
\end{equation}
The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances ($var$). Prior distributions on variances have a smaller scale than priors on standard deviation ($sd$). Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself. \Revision{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well. If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as ${sd}^2$ if we formulated the prior on the variance. From this, we can derive the implied distribution $f(var)$ or $f(sd)$. Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$. To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation. For the `brms` priors, we obtain the following implied prior distributions on variances:

@@

## \Revision{Technical Possibilities are Endless}
